---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# cobin-reproduce

<!-- badges: start -->
<!-- badges: end -->

Code to reproduce figures, simulation results, and case study results from the paper 

> Anonymous authors (2025). Scalable and robust regression models for continuous proportional data. [submitted]

<!--
> Lee, C. J., Dahl, B. K., Ovaskainen, O., & Dunson, D. B. (2025). Scalable and robust regression models for continuous proportional data. [submitted]
-->

## Prerequisites

[to be added very soon after anonymizing]
The R package `cobin` contains all necessary functions for cobin and micobin regression models. 
Download `cobin_1.0.0.0.tar.gz` and install cobin R package from the source using the following command in R: 

```{r, eval = FALSE}
# install.packages("devtools")
install.packages("path_to_downloaded_file/cobin_1.0.0.0.tar.gz", repos = NULL, type="source")
```

## Codes for reproducing simulation results

### Simulation 1: consistency under model misspecification

- `Sec5_simul1/betareg_cobit.R`: code for finding beta regression MLE with cobit link; derived from betareg package
- `Sec5_simul1/datagen_consistency.R`: code for synthetic data generation. 
- `Sec5_simul1/runsim_consistency.R`: code for running simulation.

### Simulation 2: scalablity and robustness using spatial regression under model misspecificaton

There are total 4 scenarios: rho01_n250, rho01_n500, rho02_n250, rho02_n500

- `Sec5_simul2/betareg_spatial.stan`: code for spatial beta regression model with cobit link in Stan
- `Sec5_simul2/datagen_spatial.R`: code for synthetic data generation. 
- `Sec5_simul2/[scenario]/runbeta_[scenario]`: code for running spatial beta regression.
- `Sec5_simul2/[scenario]/runcobin_[scenario]`: code for running spatial cobin regression.
- `Sec5_simul2/[scenario]/runmicobin_[scenario]`: code for running spatial micobin regression.


## Codes for reproducing benthic macroinvertibrate multimetric index (MMI) analysis

### Data preparation, exploratory data analysis 

* `Sec6_mmicasestudy/mmi_lakecat_csv`: MMI and lake watershed covariate for 950 lakes 
* `Sec6_mmicasestudy/lakecat-over40000m2.csv`: lake watershed covariate for 50k+ lakes for prediction
  - See dataprocess_readme.txt for data collecting procedure from EPA website
* `Sec6_mmicasestudy/fig_mmidata.R`: code for reproducing figure 1 (MMI data and urban cover)

### Data analysis and prediction 

* `Sec6_mmicasestudy/results_main_n949`: folder containing MMI data analysis of 949 lakes, excluding one lake with MMI = 0
* `Sec6_mmicasestudy/results_supp_n947`: folder containing MMI data analysis of 947 lakes, excluding two additional lakes from 949 lakes for sensitivity analysis
* `Sec6_mmicasestudy/results_supp_n950`: folder containing MMI data analysis of 950 lakes (only micobin regression) for sensitivity analysis
* `Sec6_mmicasestudy/fig_prediction.R`: code for reproducing figure 3 (prediction)
* `Sec6_mmicasestudy/fig_qresiduals.R`: code for reproducing figure 4 (quantile residual)


## Codes for reproducing other figures

* Figure 2 (density of cobin and micobin):
  - `others/density.R`
* Figure 2 (link function and variance function comparison):
  - `others/linkft.R`
  
