---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# cobin-reproduce

<!-- badges: start -->
<!-- badges: end -->
Code to reproduce figures, simulation results, and case study results from the paper 

> Lee, C. J., Dahl, B. K., Ovaskainen, O., Dunson, D. B. (2026). Scalable and robust regression models for continuous proportional data. Journal of the American Statistical Association, in press.  

Preprint is available at https://arxiv.org/abs/2504.15269; the journal version will be updated soon. 

## Prerequisites

The R package `cobin` contains all necessary functions for cobin and micobin regression models. 
All the analyses were based on version 1.0.0.1. Install the cobin R package from CRAN:

```{r, eval = FALSE}
install.packages("cobin")
#devtools::install_github("changwoo-lee/cobin") # development version
```

See cobin_1.0.0.1.pdf for the package documentation or type `?(function_name)` for details.

Test if the package is installed correctly, e.g. Kolmogorov-Gamma sampler:
```{r, message = F, warning = F}
library(cobin)
# theoretical mean
ekg = function(b, c){
  if(c==0) b/12 else return( b*((c/2)*(1/tanh(c/2))-1)/(c^2)) 
}

ekg(1,2) # theoretical mean
mean(cobin::rkgcpp(1000000, 1, 2)) # empirical mean

```


## Codes for reproducing simulation results

- `Simul1`: Codes for Simulation Study 1 (Robustness of MLE to model misspecification)
- `Simul2`: Codes for Simulation Study 2 (Robustness and scalability under spatial regression models)
- `Simul3`: Codes for Simulation Study 2 (Robustness to the presence of outliers)
- `MMIanalysis`: Codes for benthic macroinvertebrate multimetric index (MMI) data analysis
- `others`: Supplementary figures (density plots, link function comparisons)

### Simulation 1: Robustness of MLE to model misspecification (Section 4.1)

- `Simul1/data/`: folder for synthetic data and generation codes. 
- `Simul1/figures/`: folder containing codes for reproducing figures (Figure S.2). 
- `Simul1/betareg_cobit.R`: code for finding beta regression MLE with cobit link; derived from betareg package
- `Simul1/runsim_consistency.R`: code for running simulation under the correct mean structure.
- `Simul1/runsim_consistency_wrongmean.R`: code for running simulation under the wrong mean structure (See Supplementary Section S.5.1).

### Simulation 2: Robustness and scalability under spatial regression models (Section 4.2)

There are total 4 scenarios: rho01_n250, rho01_n500, rho02_n250, rho02_n500

- `Simul2/betareg_spatial.stan`: code for spatial beta regression model with cobit link in Stan
- `Simul2/data/`: folder for synthetic data and generation codes. 
- `Simul2/[scenario]/runbeta_[scenario]`: code for running spatial beta regression.
- `Simul2/[scenario]/runcobin_[scenario]`: code for running spatial cobin regression.
- `Simul2/[scenario]/runmicobin_[scenario]`: code for running spatial micobin regression.
- `Simul2/predict_linpred_cobit.R`: helper function for spatial prediction. 


### Simulation 3: Robustness to the presence of outliers (Section 4.3)

- `Simul3/data/`: folder for synthetic data and generation codes. 
- `Simul3/figures/`: folder containing codes for reproducing figures (Figure S.3). 
- `Simul3/betareg_fixedeffect.stan`: code for beta regression model with cobit link in Stan
- `Simul3/betarecreg_fixedeffect.stan`: code for beta rectangular regression model with cobit link in Stan
- `Simul3/run_outliersim.R`: code for running outlier robustness simulation


## Codes for reproducing benthic macroinvertibrate multimetric index (MMI) analysis (Section 5)

### Data preparation, exploratory data analysis 

* `MMIanalysis/mmi_lakecat_csv`: MMI and lake watershed covariate for 950 lakes 
* `MMIanalysis/lakecat-over40000m2.csv`: lake watershed covariate for 50k+ lakes for prediction
  - See dataprocess_readme.txt for data collecting procedure from EPA website
* `MMIanalysis/figures.R`: code for reproducing Figure 1 (MMI data and urban cover), Table 4 (Information of lakes with low MMI), Figure 3 (predicted MMI), Figure S.4 (quantile residuals), and more predicted MMI plots in supplementary
* `MMIanalysis/make_tabe.R`: code for reproducing Table 5 and more results in supplementary

### Data analysis 

* `MMIanalysis/n949`: folder containing MMI data analysis of 949 lakes, excluding one lake with MMI = 0
* `MMIanalysis/n947`: folder containing MMI data analysis of 947 lakes, excluding two additional lakes from 949 lakes
* `MMIanalysis/n950`: folder containing MMI data analysis of 950 lakes (only beta rectangular and micobin regression)


## Codes for reproducing other results

* Figure 2 (density of cobin and micobin):
  - `others/fig_density.R`
* Figure S.1 (link function and variance function comparison):
  - `others/fig_linkft.R`
* Figure S.7 and S.8 (mixing under boundary-concentrated setting):
  - `others/boundary_concentrate.R`
